---
title: "R Notebook"
output: html_notebook
---

```{r}
source("./scripts/load_dependencies.R")
```

```{r}
hp_train <- fread("./data/house_price_train.csv")
summary(hp_train)
```

We want to convert the date into individual features to split it from a very large factor variable into numeric features.

Additionally, the ID will not help us in making predictions so we will drop it.

While ZIP codes are probably an important and relevant thing to consider in prediction, there are so many of them (`r length(unique(hp_train$zipcode))`) that it makes it infeasible to 


```{r}
pp <- c(create_date_features, rm_zip, rm_id)

for (p_step in pp) {
  hp_train <- p_step(hp_train)
}

summary(hp_train)
```

Feature Creation


```{r}

original_features <- names(hp_train)

fc <- c(yrs_since_renovated, sum_grades, bed_bath_ratio, size_yard, cluster_loc)

for (create_step in fc) {
  hp_train <- create_step()
}

summary(hp_train[,setdiff(names(hp_train), original_features)])


```

Feature Selection


```{r}
final_model <- readRDS("./models/rf_model_gridsearch.rds")

final_model$coefnames
```

Cross Validation

```{r}
final_model$results

final_model$results[final_model$results$MAPE==min(final_model$results$MAPE),]
```

```{r}

```

Note: just a copy of "make_predictions.R"

```{r}
pp <- c(create_date_features, rm_zip, rm_id)
fc <- c(yrs_since_renovated, sum_grades, bed_bath_ratio, size_yard, cluster_loc)


make_predictions <- function(df, preprocessing, feature_creation, train_obj) {
  df <- data.table(df)
  
  for(f in c(preprocessing,feature_creation)) {
    df <- f(df)
  }
  
  print(names(df))
  
  df <- one_hot(df)[,train_obj$coefnames,with=F]
  
  return(predict(train_obj$finalModel, df))
}



preds <- make_predictions(hp_test, pp, fc, rf_model_gridsearch)

guesses <- data.table(id=hp_test$id, pred=preds$predictions)
fwrite(guesses, "./data/guesses.csv")
```

